---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# landmarkregUtils

<!-- badges: start -->
<!-- badges: end -->

The goal of landmarkregUtils is to provide a flexible version of the `fda::landmarkreg` function. Essentially, the computation of time warping functions $h(t)$ and their use to register curves are decoupled. 
Wrapper functions allow to either apply landmark registration directly on curve samples, thus hiding the `fda` details to the user, or on `fd` objects. In the latter case, the log rate of the warping function, i.e. $r(t) = - \log(\frac{dh}{dt})$, is also returned. 

## Installation

You can install the development version of landmarkregUtils from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("uasolo/landmarkregUtils")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(landmarkregUtils)
## basic example code
```

